Synthesising Termination and Non-Termination Proofs for Bit-Vector Programs

### 2014-07-08 papers/termination/paper.tex
Bibtex

Of the 47 benchmarks, 2 use nonlinear operations in the program (loop6 and loop11),
and 5 have nested loops (svcomp6, svcomp12, svcomp18, svcomp40, svcomp41).
{\sc Juggernaut} handles the nonlinear cases correctly and rapidly.
It solves 4 of the 5 nested loops in less than 30\,s\footnote{Although
2 of these cases timed out with our default parameters, changing one command
line flag allowed them to go through.  The benchmarks using non-default
command line flags are marked with a *.}, but times out on the 5th.

Overall, we notice that the GP component was always the fastest to find
candidate solutions, {\sc ExplicitSearch} was always fastest
to find counterexample inputs, and {\sc CBMC} was always fastest at
proving correctness.  Each of these search methods has its strengths,
and removing any one would have had a significant negative impact on
overall performance.  Approximately 90\,\% of the total
runtime was spent in synthesising new candidate solutions, which matches
our expectations -- the (non-)termination proofs are difficult to
find, but easy to check.

@@

While program termination has been extensively studied, much less research
has been conducted in the area of proving non-termination.  Gupta et
al.~dynamically enumerate lasso-shaped candidate paths for counterexamples,
and then statically prove their
feasibility~\cite{DBLP:conf/popl/GuptaHMRX08}.  Chen et al.~prove
nontermination via reduction to safety
proving~\cite{DBLP:conf/tacas/ChenCFNO14}.  Their iterative algorithm uses
counterexamples to a fixed safety property to refine an under-approximation
of a program.  In order to prove both termination and non-termination,
Harris et al.~compose several program analyses (termination provers for
multi-path loops, non-termination provers for cycles, and global safety
provers)~\cite{DBLP:conf/sas/HarrisLNR10}.  We propose a uniform treatment
of termination and non-termination by formulating a generalised second-order
formula whose solution is a proof of one of them.

>>>>

Of the 47 benchmarks, 2 use nonlinear operations in the program (loop6 and loop11),
and 5 have nested loops (svcomp6, svcomp12, svcomp18, svcomp40, svcomp41).
{\sc Juggernaut} handles the nonlinear cases correctly and rapidly.
It solves 4 of the 5 nested loops in less than 30s\footnote{Although
2 of these cases timed out with our default parameters, changing one command
line flag allowed them to go through.  The benchmarks using non-default
command line flags are marked with a *.}, but times out on the 5th.

Overall, we notice that the GP component was always the fastest to find
candidate solutions, {\sc ExplicitSearch} was always fastest
to find counterexample inputs, and {\sc CBMC} was always fastest at
proving correctness.  Each of these search methods has its strengths,
and removing any one would have had a significant negative impact on
overall performance.  Approximately 90\% of the total
runtime was spent in synthesising new candidate solutions, which matches
our expectations -- the (non-)termination proofs are difficult to
find, but easy to check.

@@

While program termination has been extensively studied, much less research
has been conducted in the area of proving non-termination.  Gupta et
al.~dynamically enumerate lasso-shaped candidate paths for counterexamples,
and then statically prove their
feasibility~\cite{DBLP:conf/popl/GuptaHMRX08}.  Chen et al.~prove
nontermination via reduction to safety
proving~\cite{DBLP:conf/tacas/ChenCFNO14}.  Their iterative algorithm uses
counterexamples to a fixed safety property to refine an under-approximation
of a program.  In order to prove both termination and non-termination,
Harris et al.~compose several program analyses (termination provers for
multi-path loops, non-termination provers for cycles, and global safety
provers)~\cite{DBLP:conf/sas/HarrisLNR10}.  We propose a uniform treatment
of termination and non-termination by formulating a generalised second order
formula whose solution is a proof of one of them.

\nocite{fisher2009phylogeny}
